Numerical Analysis \\ Assignment 7

### 2017-10-07 Graduate - Numerical Analysis/Assignment 7/assignment.tex
add problem 3

\problem{1}{Problem 3.28, Page 190.}
\solution{(a)}{
We know from the last two conditions (the first derivative),
$$
p'(x) = \frac{y_1'-y_0'}{x_1-x_0}(x-x_0)+y_0'.
$$
Integrate this polynomial we get
$$
p(x) = \frac{1}{2}\frac{y_1'-y_0'}{x_1-x_0}(x-x_0)^2+y_0'x+c.
$$
with boundary conditions we get $c = y_0-y_0'x_0$, then
$$
p(x) = y_0 + y_0'(x-x_0)\left(-\frac{(x-x_0)^2}{2(x_1-x_0)}+1\right)+y_1'\left(\frac{(x-x_0)^2}{2(x_1-x_0)}\right).
$$
}
\solution{(b)}{
We assume $p'(x_1) = y_1'$ is known. Then using Hermite interpolation we can get the fifth-order polynomial interpolation
$$
H(x) = \sum_{i=0}^{2}h_i(x) + \sum_{i=0}^{2}\tilde{h_i}(x),
$$
which satisfies
$$
\begin{aligned}
H(x_i) = y_i, ~0\le i\le 2, \\
H(x_i)' = y_i', ~0\le i\le 2.
\end{aligned}
$$
In order to get a fourth-order polynomial, we need to cancel out the coefficient of the fifth-order term, which means
$$
\sum_{i=0}^{2}(-2l_i'(x_i))\left(\frac{1}{\prod_{j\ne i}(x_i-x_j)}\right)^2y_i + \sum_{i=0}^{2}\left(\frac{1}{\prod_{j\ne i}(x_i-x_j)}\right)^2y_i' = 0.
$$
It shows
$$
\begin{aligned}
-2&\left(\frac{(x_0-x_1)+(x_0-x_2)}{(x_0-x_1)^3(x_0-x_2)^3}y_0+\frac{(x_1-x_0)+(x_1-x_2)}{(x_1-x_0)^3(x_1-x_2)^3}y_1+\frac{(x_2-x_0)+(x_2-x_1)}{(x_2-x_0)^3(x_2-x_1)^3}y_2\right) \\
+& \left(\frac{1}{(x_0-x_1)^2(x_0-x_2)^2}y_0'+\frac{1}{(x_1-x_0)^2(x_1-x_2)^2}y_1'+\frac{1}{(x_2-x_0)^2(x_2-x_1)^2}\right) = 0.
\end{aligned}
$$
with $x_i = x_0+ih$, this equation becomes
$$
y_1' = \frac{3}{4h}(y_2-y_0)-\frac{1}{4}(y_0'+y_2').
$$
Then the polynomial is 
$$
H(x) = \sum_{i=0}^{2}h_i(x) + \sum_{i=0}^{2}\tilde{h_i}(x),
$$
with
$$
\begin{aligned}
&h_i(x) = (1-2l_i'(x_i)(x-x_i))(l_i(x))^2, \\
&\tilde h_i(x) = (x-x_i)(l_i(x))^2,
\end{aligned}
$$
and $l_{i}(x)$ is the Lagrange basis.
}

@@

\problem{3}{Problem 3.38, Page}

>>>>

\problem{1}{Problem 3.28, Page 190.}
\solution{(a)}{
We know from the last two conditions (the first derivative),
$$
p'(x) = \frac{y_1'-y_0'}{x_1-x_0}(x-x_0)+y_0'.
$$
Integrate this polynomial we get
$$
p(x) = \frac{1}{2}\frac{y_1'-y_0'}{x_1-x_0}(x-x_0)^2+y_0'x+c.
$$
with boundary conditions we get $c = y_0-y_0'x_0$, then
$$
p(x) = y_0 + y_0'(x-x_0)\left(-\frac{(x-x_0)^2}{2(x_1-x_0)}+1\right)+y_1'\left(\frac{(x-x_0)^2}{2(x_1-x_0)}\right).
$$
}
\solution{(b)}{
We assume $p'(x_1) = y_1'$ is known. Then using Hermite interpolation we can get the fifth-order polynomial interpolation
$$
H(x) = \sum_{i=0}^{2}h_i(x) + \sum_{i=0}^{2}\tilde{h_i}(x),
$$
which satisfies
$$
\begin{aligned}
H(x_i) = y_i, ~0\le i\le 2, \\
H(x_i)' = y_i', ~0\le i\le 2.
\end{aligned}
$$
In order to get a fourth-order polynomial, we need to cancel out the coefficient of the fifth-order term, which means
$$
\sum_{i=0}^{2}(-2l_i'(x_i))\left(\frac{1}{\prod_{j\ne i}(x_i-x_j)}\right)^2y_i + \sum_{i=0}^{2}\left(\frac{1}{\prod_{j\ne i}(x_i-x_j)}\right)^2y_i' = 0.
$$
It shows
$$
\begin{aligned}
-2&\left(\frac{(x_0-x_1)+(x_0-x_2)}{(x_0-x_1)^3(x_0-x_2)^3}y_0+\frac{(x_1-x_0)+(x_1-x_2)}{(x_1-x_0)^3(x_1-x_2)^3}y_1+\frac{(x_2-x_0)+(x_2-x_1)}{(x_2-x_0)^3(x_2-x_1)^3}y_2\right) \\
+& \left(\frac{1}{(x_0-x_1)^2(x_0-x_2)^2}y_0'+\frac{1}{(x_1-x_0)^2(x_1-x_2)^2}y_1'+\frac{1}{(x_2-x_0)^2(x_2-x_1)^2}\right) = 0.
\end{aligned}
$$
with $x_i = x_0+ih$, this equation becomes
$$
y_1' = \frac{3}{4h}(y_2-y_0)-\frac{1}{4}(y_0'+y_2').
$$
Then the polynomial is 
$$
H(x) = \sum_{i=0}^{2}h_i(x)y_i + \sum_{i=0}^{2}\tilde{h_i}(x)y'_i,
$$
with
$$
\begin{aligned}
&h_i(x) = (1-2l_i'(x_i)(x-x_i))(l_i(x))^2, \\
&\tilde h_i(x) = (x-x_i)(l_i(x))^2,
\end{aligned}
$$
and $l_{i}(x)$ is the Lagrange basis.
}

@@

\problem{3}{Problem 3.38, Page}
\solution{Solution}{
Let $k(x) = \hat{s}(x)-g(x)$, then
$$
\begin{aligned}
\int_{x_0}^{x_n}|g''(x)|^2dx &= \int_{x_0}^{x_n}|\hat{s}''(x)-k''(x)|^2dx \\
&= \int_{x_0}^{x_n}|\hat{s}''(x)|^2dx-2\int_{x_0}^{x_n}\hat{s}''(x)k''(x)dx+\int_{x_0}^{x_n}|k''(x)|^2dx.
\end{aligned}
$$
By integration by parts, 
$$
\left.\int_{x_0}^{x_n}\hat{s}''(x)k''(x)dx = \hat{s}''(x)k'(x)\right|_{x_0}^{x_n} - \int_{x_0}^{x_n}\hat{s}'''(x)k'(x)dx
$$
We know from the boundary conditions, 
$$
\hat{s}''(x_0) = \hat{s}''(x_n) = 0,
$$
thus the first term equals to $0$. Besides, since $\hat{s}(x)$ is a cubic polynomial, $\hat{s}'''(x)$ is a constant, denoted by $c$. Then 
$$
\left.\int_{x_0}^{x_n}\hat{s}'''(x)k'(x)dx = ck(x)\right|_{x_0}^{x_n} = c(k(x_n)-k(x_0)).
$$
We know from intepolation conditions that
$$
k(x_i) = \hat{s}(x_i) - g(x_i) = 0, ~0\le i\le n
$$
Thus the second term also equals to $0$. Then we know
$$
\int_{x_0}^{x_n}|g''(x)|^2dx = \int_{x_0}^{x_n}|\hat{s}''(x)|^2dx+\int_{x_0}^{x_n}|\hat{s}''(x)-g''(x)|^2dx \ge \int_{x_0}^{x_n}|\hat{s}''(x)|^2dx.
$$
}
